


```{r,setOpts,include=FALSE}
opts_chunk$set(warnings=FALSE,echo=FALSE)
# n.P <- 1; n.M <- 1; n.Surv <- 1
``` 


# `r sapply( JSON$Survival$Endpoints  , with , Label)[n.Surv]` survival and `r sapply( JSON$Survival$Predictors , with , Label)[n.P]`

`r sapply( JSON$Survival$Endpoints , with , Label )[n.Surv]` is observed in `r sum(D$E0 , na.rm = TRUE )` of `r dim(D)[n.Surv]` patients (`r round(100*sum(D$E0 , na.rm = TRUE )/dim(D)[n.Surv] , 1)`%) with a median follow-up time of `r median(D[ , paste("T",n.Surv-1,sep="")],na.rm=TRUE)` days.


## Kaplan-Meyer plot

Kaplan-Meyer estimates of `r sapply( JSON$Survival$Endpoints , with , Label )[n.Surv]` survival over quartiles of `r sapply( JSON$Survival$Predictors , with , Label )[n.P]`.
              
```{r,KMplotcP,fig.width=6,fig.height=6,out.width="600pt"}

pal <-  c("#1B9E7788","#D95F0288","#7570B388","#E7298A88")
F <- as.formula( paste( "Surv" , n.Surv-1 , "~cP" , n.P-1 , sep = ""))
ThisSurvFit <- survfit( F , data = D )

atY <- seq(0,1,0.1)
atY <- atY[ which( 10*atY >= trunc( 10*min(ThisSurvFit$surv))) ]

plot(
    ThisSurvFit ,
    ylim = range(atY) ,
    lwd = 3 ,
    xlab = "Days" ,
    ylab = paste( "Percent" , sapply( JSON$Survival$Endpoints , with , ShortLabel )[n.Surv] , "survival" ) ,
    col = pal ,
    axes = FALSE ,
    bty = "n" ,
    mark.time = FALSE
    )

axis( 1 ) 
axis( 2 , at = atY , label = 100*atY )

legend(
    "topright" ,
    bty = "n" ,
    lwd = 5 ,
    col = pal ,
    title = sapply( JSON$Survival$Predictors , with , ShortLabel )[n.P] ,
    legend = paste( "Q",1:4,sep="" ))

``` 



## Cox-Regression: `r sapply( JSON$Survival$Endpoints, with , ShortLabel)[n.Surv]` ~ `r sapply( JSON$Survival$Predictors, with , ShortLabel)[n.P]`
  

  
```{r,TableSurvivalPredictor,results='asis'}

TableSurvivalPredictor <- rbind(
    cbind(
        OutputHR( alevel = 0 , ptrans = "z" ) ,
        OutputHR( alevel = 1 , ptrans = "z" ) ,
        OutputHR( alevel = 2 , ptrans = "z" )) ,
    cbind(
        OutputHR( alevel = 0 , ptrans = "zlog" ) ,
        OutputHR( alevel = 1 , ptrans = "zlog" ) ,
        OutputHR( alevel = 2 , ptrans = "zlog" )) ,
    cbind(
        OutputHR( alevel = 0 , ptrans = "n" ) ,
        OutputHR( alevel = 1 , ptrans = "n" ) ,
        OutputHR( alevel = 2 , ptrans = "n" )))

rownames( TableSurvivalPredictor ) <- paste(
    c("Normalized","Normalized logarithm of","Trend over Quartiles of") ,
    sapply( JSON$Survival$Predictors , with , ShortLabel )[n.P])

htmlTable(
    TableSurvivalPredictor ,
    cgroup = c("Unadjusted" , "Adjustment 1" , "Adjustment 2") ,
    n.cgroup = c(3,3,3) ,
    caption = paste(
        "Hazard ratios for association between" ,
        sapply( JSON$Survival$Endpoints , with , Label )[n.Surv] ,
        "and" , 
        sapply( JSON$Survival$Predictors , with , Label )[n.P] ,
        "estimated with Cox proportional hazards models; adjustment 1 for" , 
        paste0( sapply( JSON$Survival$Adjustments , with , Label )[sapply( JSON$Survival$Adjustments , with , Level )==1] , collapse = ', ') ,
        "; adjustment 2 additionally for" , 
        paste0( sapply( JSON$Survival$Adjustments , with , Label )[sapply( JSON$Survival$Adjustments , with , Level )==2] , collapse = ', ') ,
        "."))

``` 

