


## 3d-GAM

Surface spline of `r sapply( JSON$Endpoints , with , Label )[n.Surv]` incidence by empirical percentiles of `r sapply( JSON$Predictors , with , Label )[n.P]` and `r sapply( JSON$Modifiers , with , Label )[n.M]` estimated with a Generalized Additve Model.

```{r,VisGAMq,fig.width=6,fig.height=6,out.width="600pt"}

F <- as.formula( paste( "E" , n.Surv-1 , "~s(qP" , n.P-1 , ",qM" , n.M-1 , ")" , sep = "" ))

GAM <- gam(
    formula = F ,
    data = D ,
    family = "binomial")

vis.gam(
    x = GAM ,
    n.grid = 50 ,
    color = "terrain" ,
    type = "response" ,
    xlab = sapply( JSON$Predictors , with , Label )[n.P] ,
    ylab = sapply( JSON$Modifiers , with , Label )[n.M] ,
    zlab = "Incidence" ,
    phi = 0 ,
    theta = -20 ,
    ticktype = "detailed"
)
    
``` 
    


## 3d-GAM with SE

Surface spline of `r sapply( JSON$Endpoints , with , Label )[n.Surv]` incidence by empirical percentiles of `r sapply( JSON$Predictors , with , Label )[n.P]` and `r sapply( JSON$Modifiers , with , Label )[n.M]` estimated with a Generalized Additve Model; coloured grids depict 95\% confidence levels.
      
```{r,VisGAMqSE,fig.width=6,fig.height=6,out.width="600pt"}

vis.gam(
    x = GAM ,
    n.grid = 50 ,
    se = TRUE ,
    type = "response" ,
    xlab = sapply( JSON$Predictors , with , Label )[1] ,
    ylab = sapply( JSON$Modifiers , with , Label )[1] ,
    zlab = "Incidence" ,
    phi = 0 ,
    theta = -20 ,
    ticktype = "detailed"
)
    
``` 
