

# Patient characteristics at baseline by `r sapply( JSON$Predictors , with , Label )[n.P]`


```{r,PCgetAllVars,include=FALSE}

opts_chunk$set(warnings=FALSE,echo=FALSE,errors=FALSE)

Table1Types <- sapply(c(JSON$Modifiers,JSON$Adjustments,JSON$PatientCharacteristics) , with , Type )

Table1Vars <- c(Ms,As,PCs)
Table1NumVars <- Table1Vars[which(Table1Types == "numerical")]
Table1CatVars <- Table1Vars[which(Table1Types == "categorical")]

Table1Labels <- sapply(c(JSON$Modifiers,JSON$Adjustments,JSON$PatientCharacteristics) , with , Label )
Table1NumLabels <- Table1Labels[which(Table1Types == "numerical")]
Table1CatLabels <- Table1Labels[which(Table1Types == "categorical")]

``` 



```{r,LaTeXproportions,results='as.is'}

Table1Proportions <- matrix(
    unlist(
        lapply(
            X = paste( Table1CatVars ) ,
            FUN = ProportionsOverGroups ,
            groups = paste("c4P" , n.P-1 , sep = "" ))
        ) ,
    ncol = 9 ,
    byrow = TRUE
)

colnames(Table1Proportions) <- c("%","N","%","N","%","N","%","N","P")
rownames(Table1Proportions) <- Table1CatLabels

kable(
    x = Table1Proportions ,
    format = "html" ,
    caption = paste("Patient characteristics over" , Predictors[n.P] , "quartiles for categorical variables reported as proportions (counts). Trends over quartiles where estimated with a logistic regression model.")
    )

``` 
    

    
    
Patient characteristics over `r Predictors[n.P]` quartiles for continuous variables reported as means (standard error).
Standard errors and p for trend over quartiles were estimated with general linear models.


```{r,LaTeXMeans}

Table1Means <- matrix(
    unlist(
        lapply(
            X = paste( Table1NumVars ) ,
            FUN = MeansOverGroups ,
            groups = paste("c4P" , n.P-1 , sep = "" ))
        ) ,
    ncol = 9 ,
    byrow = TRUE
)

colnames(Table1Means) <- c( "Mean","SE","Mean","SE","Mean","SE","Mean","SE","P" )
rownames(Table1Means) <- Table1NumLabels

kable(Table1Means)

```
    

Patient characteristics over `r Predictors[n.P]` quartiles for continuous variables reported as medians (standard error).
Standard errors and p for trend over quartiles were estimated with median regression models.

```{r,LaTeXMedians}

Table1Medians <- matrix(
    unlist(
        lapply(
            X = paste( Table1NumVars ) ,
            FUN = MediansOverGroups ,
            groups = paste("c4P" , n.P-1 , sep = "" ))
        ) ,
    ncol = 9 ,
    byrow = TRUE
)

colnames(Table1Medians) <- c( "Median","SE","Median","SE","Median","SE","Median","SE","P" )
rownames(Table1Medians) <- Table1NumLabels

kable(Table1Medians)


``` 
