


```{r,setOpts,include=FALSE}
opts_chunk$set(warnings=FALSE,echo=FALSE)
# n.P <- 2; n.M <- 1; n.Surv <- 1
``` 


# Interaction: `r sapply( JSON$Survival$Predictors , with , ShortLabel )[n.P]` x `r sapply( JSON$Survival$Modifiers , with ,ShortLabel )[n.M]`
## Endpoint: `r sapply( JSON$Survival$Endpoints  , with , Label)[n.Surv]`
## Predictor: `r sapply( JSON$Survival$Predictors , with , Label)[n.P]`
## Modifier:  `r sapply( JSON$Survival$Modifiers  , with , Label)[n.M]`


## `r sapply( JSON$Survival$Endpoints , with , ShortLabel )[n.Surv]` and `r sapply( JSON$Survival$Predictors , with , ShortLabel )[n.P]` in subgroups of high/low `r sapply( JSON$Survival$Modifiers , with , ShortLabel )[n.M]`

      
```{r,TabCoxSubgroup,results='asis'}

TabCoxSubgroup <- rbind(
        PrintCoxSubgroupTable( alevel = 0 , ptrans = "z" ) ,
        PrintCoxSubgroupTable( alevel = 1 , ptrans = "z" ) ,
        PrintCoxSubgroupTable( alevel = 2 , ptrans = "z" ) ,
        PrintCoxSubgroupTable( alevel = 0 , ptrans = "zlog" ) ,
        PrintCoxSubgroupTable( alevel = 1 , ptrans = "zlog" ) ,
        PrintCoxSubgroupTable( alevel = 2 , ptrans = "zlog" ) ,
        PrintCoxSubgroupTable( alevel = 0 , ptrans = "n" ) ,
        PrintCoxSubgroupTable( alevel = 1 , ptrans = "n" ) ,
        PrintCoxSubgroupTable( alevel = 2 , ptrans = "n" ))
        
htmlTable(
    TabCoxSubgroup ,
    rnames = rep( c("Unadjusted","Adjustment 1","Adjustment 2") , 3 ) ,
    rgroup = paste(
        c("Per SD increase of","Per SD increase of log-transformed","Trend over quartiles of") ,
        sapply( JSON$Survival$Predictors , with , Label )[n.P]) ,
    n.rgroup = c(3,3,3) ,
    cgroup = c( 
        paste( "Low" , sapply( JSON$Survival$Modifiers , with , ShortLabel )[n.M] ) ,
        paste( "High" , sapply( JSON$Survival$Modifiers , with , ShortLabel )[n.M] ) ,
        "Between group" ) ,
    n.cgroup = c(2,2,1) ,
    caption = paste(
        "Hazard ratios for association between" ,
        sapply( JSON$Survival$Endpoints , with , Label )[n.Surv] ,
        "and" ,
        sapply( JSON$Survival$Predictors , with , Label )[n.P] ,
        "in subgroups of high vs. low levels of" ,
        sapply( JSON$Survival$Modifiers , with , Label )[n.M] ,
        "estimated with Cox proportional hazards models; adjustment 1 for" ,
        paste0( sapply( JSON$Survival$Adjustments , with , Label )[sapply( JSON$Survival$Adjustments , with , Level )==1] , collapse = ', ') ,
        "; adjustment 2 additionally  for" ,
        paste0( sapply( JSON$Survival$Adjustments , with , Label )[sapply( JSON$Survival$Adjustments , with , Level )==2] , collapse = ', ') ,
        ". The between group p-value is estimated from the interaction term with the effect modifier" ,
        sapply( JSON$Survival$Modifiers , with , Label )[n.M] ,
        "divided into high/low levels according to the population median of" ,
        signif( median( D[ , paste( "M" , n.M-1 , sep = "" ) ] , na.rm = TRUE ),3) ,
        sapply( JSON$Survival$Modifiers , with , unit )[n.M]
        )
    )
        
        
``` 



## `r sapply( JSON$Survival$Endpoints , with , ShortLabel )[n.Surv]` and `r sapply( JSON$Survival$Predictors , with , ShortLabel )[n.P]` and interaction with high/low `r sapply( JSON$Survival$Modifiers , with , ShortLabel )[n.M]`

      
```{r,TabCoxInteraction,results='asis'}

TabCoxInteraction <- rbind(
    cbind(
        OutputHR( alevel = 0 , ptrans = "z" , mtrans = "c" ) ,
        OutputHR( alevel = 1 , ptrans = "z" , mtrans = "c" ) ,
        OutputHR( alevel = 2 , ptrans = "z" , mtrans = "c" )) ,
                
    cbind(
        OutputHR( alevel = 0 , ptrans = "zlog" , mtrans = "c" ) ,
        OutputHR( alevel = 1 , ptrans = "zlog" , mtrans = "c" ) ,
        OutputHR( alevel = 2 , ptrans = "zlog" , mtrans = "c" )) ,

    cbind(
        OutputHR( alevel = 0 , ptrans = "n" , mtrans = "c" ) ,
        OutputHR( alevel = 1 , ptrans = "n" , mtrans = "c" ) ,
        OutputHR( alevel = 2 , ptrans = "n" , mtrans = "c" )))


htmlTable(
    TabCoxInteraction ,
    rnames = c(
         paste( "Per SD increase" , sapply( JSON$Survival$Predictors , with , Label )[n.P] ) ,
         paste( "High vs. low" , sapply( JSON$Survival$Modifiers , with , Label )[n.M] ) ,
         "Multiplicative effect modification" ,
         paste( "Per SD increase of log-transformed" , sapply( JSON$Survival$Predictors , with , Label )[n.P] ) ,
         paste( "High vs. low" , sapply( JSON$Survival$Modifiers , with , Label )[n.M] ) ,
         "Multiplicative effect modification" ,
         paste( "Per quartile increase of" , sapply( JSON$Survival$Predictors , with , Label )[n.P] ) ,
         paste( "High vs. low" , sapply( JSON$Survival$Modifiers , with , Label )[n.M] ) ,
         "Multiplicative effect modification") ,
    rgroup = c("Normalized" , "Normalized logarithmic" , "Trend over quartiles" ) ,
    n.rgroup = c(3,3,3) ,
    cgroup = c("Unadjusted","Adjustment 1","Adjustment 2") ,
    n.cgroup = c(3,3,3) ,
    caption = paste(
        "Hazard ratios for association between" ,
        sapply( JSON$Survival$Endpoints , with , Label )[1] ,
        "and" ,
        sapply( JSON$Survival$Predictors , with , Label )[1] ,
        "and effect modification by high vs low levels of" ,
        sapply( JSON$Survival$Modifiers , with , Label )[1] ,
        "( divided into high/low levels according to the population median of" ,
        signif( median( D[ , paste( "M" , n.M-1 , sep = "" ) ] , na.rm = TRUE ),3) ,
        ")  estimated with Cox proportional hazards models; adjustment 1 for" ,
        paste0( sapply( JSON$Survival$Adjustments , with , Label )[sapply( JSON$Survival$Adjustments , with , Level )==1] , collapse = ', ') ,
        "; adjustment 2 additionally  for" ,
        paste0( sapply( JSON$Survival$Adjustments , with , Label )[sapply( JSON$Survival$Adjustments , with , Level )==2] , collapse = ', ') ,
        "."
        ))

``` 



## GAM plot percentiles (ranked)

Smoothing spline estimates of `r sapply( JSON$Survival$Endpoints , with , Label )[n.Surv]` hazard by percentiles of `r sapply( JSON$Survival$Predictors , with , Label )[n.P]` in subgroups of high vs. low levels of `r sapply( JSON$Survival$Modifiers , with , Label )[n.M]` estimated with Cox proportional hazards models adjusted for `r paste0( sapply( JSON$Survival$Adjustments , with , Label )[sapply( JSON$Survival$Adjustments , with , Level )==1] , collapse = ', ')`. The effect modifier `r sapply( JSON$Survival$Modifiers , with , Label )[n.M]` is divided into high/low levels according to the population median of `r signif( median( D[ , paste( "M" , n.M-1 , sep = "" ) ] , na.rm = TRUE ),3)`.
              
```{r,CoxSmoothSubgroupsq,fig.width=6,fig.height=6,out.width="600pt"}

PlotCoxSmoothSubgroups(
    ptrans = "q" ,
    alevel = 1 ,
    xlab = paste("Percentiles of" , sapply( JSON$Survival$Predictors , with , Label )[n.P])
)

``` 


## GAM plot normalized logarithm

Smoothing spline estimates of `r sapply( JSON$Survival$Endpoints , with , Label )[n.Surv]` hazard by SD of log-transformed `r sapply( JSON$Survival$Predictors , with , Label )[n.P]` in subgroups of high vs. low levels of `r sapply( JSON$Survival$Modifiers , with , Label )[n.M]` estimated with Cox proportional hazards models adjusted for `r paste0( sapply( JSON$Survival$Adjustments , with , Label )[sapply( JSON$Survival$Adjustments , with , Level )==1] , collapse = ', ')`. The effect modifier `r sapply( JSON$Survival$Modifiers , with , Label )[n.M]` is divided into high/low levels according to the population median of `r signif( median( D[ , paste( "M" , n.M-1 , sep = "" ) ] , na.rm = TRUE ),3)`.

```{r,CoxSmoothSubgroupszlog,fig.width=6,fig.height=6,out.width="600pt"}

PlotCoxSmoothSubgroups(
    ptrans = "zlog" ,
    alevel = 1 ,
    xlab = paste("SD of log-transformed" , sapply( JSON$Survival$Predictors , with , Label )[n.P])
)

``` 
